---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
#Load necessary libraries for analyses
```{r}
library(ggthemes)
library(RColorBrewer)
library(extrafont)
library(grid)
library(here)
library(tidyverse)
library(ggrepel)
library(cowplot)
library(car)
library(vioplot)
library(ggplot2)
library(ggsci)
library(ggsignif)
library(tidyverse)
library(ggstatsplot)
library(FSA)
```

```{r}
setwd("/Users/ashleydye/Documents/THESIS_Calls and Genes Project/Final_data_analysis/")
frog_calls <- read.csv("final_call_analysis.csv", header=TRUE, sep=",")
frog_calls
```
#Check relationship between call traits and temperature
```{r}
#Dominant frequency
plot(frog_calls$Temperature, frog_calls$Dominant.frequency, xlab="Temperature", ylab="Dominant Frequency")
abline(lm(frog_calls$Dominant.frequency~frog_calls$Temperature))
cor.test(frog_calls$Dominant.frequency, frog_calls$Temperature, method = "pearson")

#Pulse rate
plot(frog_calls$Temperature, frog_calls$Pulse.rate, xlab="Temperature", ylab="Pulse rate")
abline(lm(frog_calls$Pulse.rate~frog_calls$Temperature))
cor.test(frog_calls$Pulse.rate, frog_calls$Temperature, method = "pearson")

#Note length
plot(frog_calls$Temperature, frog_calls$Note.length, xlab="Temperature", ylab="Note length")
abline(lm(frog_calls$Note.length~frog_calls$Temperature))
cor.test(frog_calls$Note.length, frog_calls$Temperature, method = "pearson")
```
#Regression with Trait and Temp to correct to specific temperature
```{r}
lm(data=frog_calls, Dominant.frequency~Temperature)

lm(data=frog_calls, Pulse.rate~Temperature)

lm(data=frog_calls, Note.length~Temperature)
```
#Check relationship between call characteristics and SVL (DO AFTER TEMP CORRECTION)
```{r}
#Dominant frequency is no longer correlated with SVL
plot(frog_calls$SVL, frog_calls$DF.TEMP, xlab="SVL", ylab="Dominant frequency")
abline(lm(frog_calls$DF.TEMP~frog_calls$SVL))
cor.test(frog_calls$DF.TEMP, frog_calls$SVL, method = "pearson")

#Pulse rate
plot(frog_calls$SVL, frog_calls$PR.TEMP, xlab="SVL", ylab="Pulse rate")
abline(lm(frog_calls$PR.TEMP~frog_calls$SVL))
cor.test(frog_calls$PR.TEMP, frog_calls$SVL, method = "pearson")

#Note length
plot(frog_calls$SVL, frog_calls$NL.TEMP, xlab="SVL", ylab="Note length")
abline(lm(frog_calls$NL.TEMP~frog_calls$SVL))
cor.test(frog_calls$NL.TEMP, frog_calls$SVL, method = "pearson")
```
#Make new spreadsheet with SR for analysis
```{r}
new_frog_calls <- frog_calls[-c(63,64),]
new_frog_calls
```
#Run Shapiro-Wilkes test for normality in sites and morphs
```{r}
#Dominant frequency is NOT normally distributed
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Achinamisa"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Callanayacu"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Chipesa"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Varadero Striped"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Vaquero"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Red-Headed Varadero"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Chazuta"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="San Jose"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Site=="Ricardo Palma"])

#Santa Rosa excluded for sample size of 2
#shapiro.test(frog_calls$DF.TEMPl[frog_calls$Site=="Santa Rosa"])

shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Striped"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Intermediate"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Banded"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Spotted"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Varadero Red-headed"])
shapiro.test(frog_calls$DF.TEMP[frog_calls$Morph=="Varadero Striped"])
```
```{r}
#Pulse rate IS NORMALLY DISTRIBUTED
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Achinamisa"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Callanayacu"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Chipesa"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Varadero Striped"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Vaquero"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Red-Headed Varadero"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Chazuta"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="San Jose"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Ricardo Palma"])

#Santa Rosa excluded for sample size of 2
#shapiro.test(frog_calls$PR.TEMP[frog_calls$Site=="Santa Rosa"])

shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Striped"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Intermediate"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Banded"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Spotted"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Varadero Red-headed"])
shapiro.test(frog_calls$PR.TEMP[frog_calls$Morph=="Varadero Striped"])

#shapiro.test(frog_calls$PR.TEMP)
```
```{r}
#Note length is NOT normally distributed
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Achinamisa"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Callanayacu"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Chipesa"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Varadero Striped"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Vaquero"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Red-Headed Varadero"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Chazuta"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="San Jose"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Ricardo Palma"])

#Santa Rosa excluded for sample size of 2
#shapiro.test(frog_calls$NL.TEMP[frog_calls$Site=="Santa Rosa"])

shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Striped"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Intermediate"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Banded"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Spotted"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Varadero Red-headed"])
shapiro.test(frog_calls$NL.TEMP[frog_calls$Morph=="Varadero Striped"])
```
#Levene's test of equal variances
```{r}
#All data has equal variances
leveneTest(DF.TEMP~ Morph, data = frog_calls)
leveneTest(PR.TEMP ~ Morph, data = frog_calls)
leveneTest(NL.TEMP ~ Morph, data = frog_calls)

leveneTest(DF.TEMP ~ Site, data = frog_calls)
leveneTest(PR.TEMP ~ Site, data = frog_calls)
leveneTest(NL.TEMP ~ Site, data = frog_calls)
```
#Calculate mean and SD for call traits
```{r}
group_mean_df_morph<- aggregate(x= frog_calls$DF.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = mean)
print(group_mean_df_morph)

group_sd_df_morph<- aggregate(x= frog_calls$DF.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = sd)
print(group_sd_df_morph)
```
```{r}
group_mean_pr_morph<- aggregate(x= frog_calls$PR.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = mean)
print(group_mean_pr_morph)

group_sd_pr_morph<- aggregate(x= frog_calls$PR.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = sd)
print(group_sd_pr_morph)
```
```{r}
group_mean_nl_morph<- aggregate(x= frog_calls$NL.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = mean)
print(group_mean_nl_morph)

group_sd_nl_morph<- aggregate(x= frog_calls$NL.TEMP,
                      # Specify group indicator
                      by = list(frog_calls$Morph),      
                      # Specify function (i.e. mean)
                      FUN = sd)
print(group_sd_nl_morph)
```
#Run Kruskal-Wallis non-parametric test for differences in means for dominant frequency and note length
```{r}
#Dominant frequency by morph
kruskal.test(DF.TEMP ~ Morph, data = frog_calls)
dunnTest(DF.TEMP ~ Morph, data = frog_calls, method = "bonferroni")
```
```{r}
#Dominant frequency by site
kruskal.test(DF.TEMP ~ Site, data = new_frog_calls)
dunnTest(DF.TEMP ~ Site, data = new_frog_calls, method = "bonferroni")
```
```{r}
#Note length by morph
kruskal.test(NL.TEMP ~ Morph, data = frog_calls)
dunnTest(NL.TEMP ~ Morph, data = frog_calls, method = "bonferroni")
```
```{r}
#Note length by Site
kruskal.test(NL.TEMP ~ Site, data = new_frog_calls)
dunnTest(NL.TEMP ~ Site, data = new_frog_calls, method = "bonferroni")
```
#Run ANOVA with TukeyHSD for pulse rate (normally distributed, no difference in variance between groups)
```{r}
#Pulse rate by morph
summary(pr.anova.morph <- aov(frog_calls$PR.TEMP~frog_calls$Morph))

pr.tukey.morph <- TukeyHSD(pr.anova.morph)
pr.tukey.morph
```
```{r}
#Pulse rate by site
summary(pr.anova.site <- aov(new_frog_calls$PR.TEMP~new_frog_calls$Site))

pr.tukey.site <- TukeyHSD(pr.anova.site)
pr.tukey.site
```
#Plots
```{r}
#Dominant frequency morph
df_morph <- ggplot(frog_calls, aes(newlabels, DF.TEMP,))
df<-df_morph +geom_violin(size=2)+ 
  stat_summary(fun = "mean",
               geom = "crossbar",
               width = 0.5,
               color = "black")+
  #scale_color_brewer(palette = "Reds", direction = 1)+
  geom_jitter(cex=5,colour="black",position = position_jitter(seed = 1, width = 0.1)) +
  labs(
  x="Color Morph",
  y="Dominant Frequency\n(Hz)"
)+theme_classic()+theme(legend.position = "none")  +  scale_x_discrete(limits=c("Spotted","Banded","Intermediate","Striped","Varadero\nStriped", "Varadero\nRed-headed")) + theme(text=element_text(size=40))
ggsave("df_morph_newlab.png", width =20, height = 9)
df
```
```{r}
#Note length site
nl_morph <- ggplot(frog_calls, aes(newlabels, NL.TEMP))
nl <- nl_morph+geom_violin(size=2)+ 
  stat_summary(fun = "mean",
               geom = "crossbar",
               width = 0.5,
               color = "black")+
  #scale_color_brewer(palette="Blues", direction = -1)+
  labs(
  x="Color Morph",
  y="Note Length\n(s)"
)+theme_classic()+theme(legend.position = "none") + geom_jitter(cex=5,colour="black",position = position_jitter(seed = 1, width = 0.1)) + scale_x_discrete(limits=c("Spotted","Banded","Intermediate","Striped","Varadero\nStriped", "Varadero\nRed-headed")) + theme(text=element_text(size=40))
ggsave("nl_morph_new.png", width =20, height = 9)
nl
```
```{r}
frog_calls$newlabels=str_wrap(frog_calls$Morph,width = 9)
#frog_calls
#Pulse rate morph
pr_morph <- ggplot(frog_calls, aes(newlabels, PR.TEMP))
pr <- pr_morph+geom_violin(size=2)+ 
  stat_summary(fun = "mean",
               geom = "crossbar",
               width = 0.5,
               color = "black")+
  #scale_color_brewer(palette = "Reds", direction = 1)+
  labs(
  x="Color Morph",
  y="Pulse Rate\n(pulses/s)"
)+theme_classic()+theme(legend.position = "none") + geom_jitter(cex=5,colour="black",position = position_jitter(seed = 1, width = 0.1)) + scale_x_discrete(limits=c("Spotted","Banded","Intermediate","Striped","Varadero\nStriped", "Varadero\nRed-headed")) + theme(text=element_text(size=40))
ggsave("pr_morph_new.png", width =20, height = 9)
pr
```

```{r}
citation()
version
```
##Mantel Tests Attempt
```{r}
#install.packages("vegan")
library(vegan)
#install.packages("geosphere")
library(geosphere)
```

```{r}
df = read.csv("final_call_analysis.csv", header= TRUE)
#call trait
calls = df[,16:ncol(df)]
freq=df$DF.TEMP
pulserate=df$PR.TEMP
notelength=df$NL.TEMP

#environmental vector
temp = df$Temperature

#longitude and latitude 
geo = data.frame(df$Latitude, df$Longitude)
```

```{r}
#abundance data frame - bray curtis dissimilarity
dist.call = dist(calls, method = "euclidean")
dist.df <- dist(freq, method = "euclidean")
dist.pr <- dist(pulserate, method = "euclidean")
dist.nl <- dist(notelength, method = "euclidean")


#environmental vector - euclidean distance
dist.temp = dist(temp, method = "euclidean")

#geographic data frame - haversine distance 
d.geo = distm(geo, fun = distHaversine)
dist.geo = as.dist(d.geo)
```

```{r}
#temp vs call
call_temp = mantel(dist.call, dist.temp, method = "spearman", permutations = 9999, na.rm = TRUE)
call_temp

#distance vs call
call_geo  = mantel(dist.call, dist.geo, method = "spearman", permutations = 9999, na.rm = TRUE)
call_geo

##Weak positive correlation between call traits and distance r=0.1059,p=0.0128
```

```{r}
#df vs geographic 
df_geo  = mantel(dist.df, dist.geo, method = "spearman", permutations = 10000, na.rm = TRUE)
df_geo

#pr vs geographic 
pr_geo  = mantel(dist.pr, dist.geo, method = "spearman", permutations = 100000, na.rm = TRUE)
pr_geo

#nl vs geographic 
nl_geo  = mantel(dist.nl, dist.geo, method = "spearman", permutations = 1000000, na.rm = TRUE)
nl_geo
```

```{r}
#install.packages("egg")
#library(egg)
multi_figure <- ggarrange(df, pr, nl, labels = c("A","B","C"))
multi_figure
```

```{r}
sessionInfo()
```

#Plot aifm1 genotypes with dominant frequency
```{r}
#new dataframe with aifm1 genotypes and individual ID
geno <- read.csv("genotypes.csv")
geno

#join to frog calls datafram
frog_calls <- left_join(frog_calls, geno, by = "DNA.ID")
```


```{r}
display.brewer.all(colorblindFriendly = TRUE)
#plot dominant frequency by genotype
df_morph <- ggplot(frog_calls, aes(Aifm1_Genotype, DF.TEMP,))
df<-df_morph +geom_violin(size=2) + 
  stat_summary(fun = "mean",
               geom = "crossbar",
               width = 0.5,
               color = "black") +
  geom_point(aes(col=Morph),size=4,position=position_jitter(w=0.1,h=0),show.legend = T) +
  scale_color_brewer(palette="Paired", direction=-1) +
  labs(x="Genotype", y="Dominant Frequency\n(Hz)") +
  theme_classic() +
  theme(legend.position = "right") + 
  theme(legend.text=element_text(size=20)) +
  theme(text=element_text(size=30))
ggsave("df_geno.png", width =10, height = 7)
df
```
```{r}
#Dominant frequency by aifm1 genotype
kruskal.test(DF.TEMP ~ Aifm1_Genotype, data = frog_calls)
dunnTest(DF.TEMP ~ Aifm1_Genotype, data = frog_calls, method = "bonferroni")
```



